NN_Accel_Heading

### 2024-03-06 Fish Data Analysis/NN_Accel_Heading.qmd
Commit before moving accel and heading code

## Loading in the Libraries for Everything
```{r}
library(MASS)
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(lme4)
```

@@

nn_accel_heading_df <- read.csv("Data/Fish_NN_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", ")))

nn_heading_sum <- nn_accel_heading_df %>% mutate(angle_bins = round_any(Angle_to_NN,pi/4)) %>%
                                                #Limit to attraction zone
                                                #filter(Distance_to_NN <= 2) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Angle_PS, na.rm = T),
                                                          sd_turn = sd(Angle_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

nn_accel_sum <- nn_accel_heading_df %>% mutate(distance_bins = round_any(Distance_to_NN,0.5)) %>%
                                                group_by(distance_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_accel = mean(Acceleration, na.rm = T),
                                                          sd_accel = sd(Acceleration, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

@@

heading_base_lm <- lm(Angle_PS ~ Flow_Ablation_Darkness, data = nn_accel_heading_df %>% filter(Flow == "Still Water"))
heading_lin_lm <- lm(Angle_PS ~ Angle_to_NN*Flow_Ablation_Darkness, data = nn_accel_heading_df %>% filter(Flow == "Still Water"))
heading_cos_lm <- lm(Angle_PS ~ cos(Angle_to_NN)*Flow_Ablation_Darkness, data = nn_accel_heading_df %>% filter(Flow == "Still Water"))
heading_sin_lm <- lm(Angle_PS ~ sin(Angle_to_NN)*Flow_Ablation_Darkness, data = nn_accel_heading_df %>% filter(Flow == "Still Water"))

@@

Anova(heading_sin_lm)

ggplot(nn_accel_heading_df %>% filter(Flow == "Still Water"),
       aes(x = Angle_to_NN, y = Angle_PS, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  theme_classic() +
  xlab("direction of neighbour (rad)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")

AIC(heading_sum_base_lm,heading_sum_lin_lm,heading_sum_cos_lm,heading_sum_sin_lm) %>% arrange(AIC)

@@

accel_base_lm <- lm(Acceleration ~ Flow_Ablation_Darkness, data = nn_accel_heading_df %>% filter(Flow == "Still Water") %>% na.omit())
accel_lin_lm <- lm(Acceleration ~ Distance_to_NN*Flow_Ablation_Darkness, data = nn_accel_heading_df %>% filter(Flow == "Still Water") %>% na.omit())

@@

Anova(accel_lin_lm)

ggplot(nn_accel_heading_df %>% filter(Flow == "Still Water"),
       aes(x = Distance_to_NN, y = Acceleration, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ x)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  ylim(-0.2, 0.2) +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

@@

>>>>

## Loading in the Libraries for Everything
```{r}
library(MASS)
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(lme4)
library(viridis)
```

@@

nn_accel_heading_df <- read.csv("Data/Fish_NN_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           mutate(Distance_to_NN = ifelse(Relative_X < 1, Distance_to_NN*-1, Distance_to_NN))

nn_accel_heading_df_tb <- nn_accel_heading_df %>% group_by(Year,Month,Day,Trial,Individual,Tailbeat,
                                                           Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                  summarise(Angle_to_NN = mean(Angle_to_NN, na.rm = T),
                                                            Angle_PS = mean(Angle_PS, na.rm = T),
                                                            Distance_to_NN = mean(Distance_to_NN, na.rm = T),
                                                            Acceleration = mean(Acceleration, na.rm = T),
                                                            Relative_X = mean(Relative_X, na.rm = T),
                                                            Relative_Y = mean(Relative_Y, na.rm = T)) %>%
                                                  ungroup() %>%
                                                  na.omit()

nn_heading_sum <- nn_accel_heading_df_tb %>% mutate(angle_bins = round_any(Angle_to_NN,pi/4)) %>%
                                                #Limit to attraction zone
                                                #filter(Distance_to_NN <= 2) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Angle_PS, na.rm = T),
                                                          sd_turn = sd(Angle_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

nn_accel_sum <- nn_accel_heading_df_tb %>% mutate(distance_bins = round_any(Distance_to_NN,0.5)) %>%
                                                group_by(distance_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_accel = mean(Acceleration, na.rm = T),
                                                          sd_accel = sd(Acceleration, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

@@

heading_base_lm <- lm(Angle_PS ~ Flow_Ablation_Darkness, data = nn_accel_heading_df_tb %>% filter(Flow == "Still Water"))
heading_lin_lm <- lm(Angle_PS ~ Angle_to_NN*Flow_Ablation_Darkness, data = nn_accel_heading_df_tb %>% filter(Flow == "Still Water"))
heading_cos_lm <- lm(Angle_PS ~ cos(Angle_to_NN)*Flow_Ablation_Darkness, data = nn_accel_heading_df_tb %>% filter(Flow == "Still Water"))
heading_sin_lm <- lm(Angle_PS ~ sin(Angle_to_NN)*Flow_Ablation_Darkness, data = nn_accel_heading_df_tb %>% filter(Flow == "Still Water"))

@@

Anova(heading_cos_lm)

ggplot(nn_accel_heading_df_tb %>% filter(Flow == "Still Water"),
       aes(x = Angle_to_NN, y = Angle_PS, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  theme_classic() +
  ylim(-0.1,0.1)+
  xlab("direction of neighbour (rad)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")

@@

accel_base_lm <- lm(Acceleration ~ Flow_Ablation_Darkness, data = nn_accel_heading_df_tb %>% filter(Flow == "Still Water"))
accel_lin_lm <- lm(Acceleration ~ Distance_to_NN*Flow_Ablation_Darkness, data = nn_accel_heading_df_tb %>% filter(Flow == "Still Water"))

@@

Anova(accel_base_lm)

ggplot(nn_accel_heading_df_tb %>% filter(Flow == "Still Water"),
       aes(x = Distance_to_NN, y = Acceleration, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ x)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  ylim(-0.01, 0.01) +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

@@

directional_df <- nn_accel_heading_df_tb %>% mutate(round_X = round_any(Relative_X,0.5),
                                                    round_Y = round_any(Relative_Y,0.5)) %>%
                                                      group_by(round_X,round_Y,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                      summarise(mean_accel = mean(Acceleration),
                                                                mean_turn = mean(Angle_PS)) %>%
                                                      ungroup()

ggplot(directional_df %>% filter(Flow == "Still Water"),
       aes(round_X, round_Y, fill= mean_accel)) + 
  geom_tile()+
  facet_wrap(~Flow_Ablation_Darkness)+
  scale_fill_viridis(option="magma")+
  coord_fixed()+
  theme_classic()

ggplot(directional_df %>% filter(Flow == "Still Water"),
       aes(round_X, round_Y, fill= mean_turn)) + 
  geom_tile()+
  facet_wrap(~Flow_Ablation_Darkness)+
  scale_fill_viridis(option="magma")+
  coord_fixed()+
  theme_classic()
